---
title: "Check the calibration of FSR/FDR of mash"
author: "Dongyue Xie"
date: "2021-04-01"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

```{r}
source('code/simulation/check_mash_lfsr.R')
```

```{r}
R = 5
n = 500
seed = 12345
```

## example 1

```{r}
u = toeplitz(c(1,0.5,0.2,0.2,0.1))
u[3:5,] = 0
u[,3:5] = 0
u
Ulist = list(U3 = u)
```


1. Using ED + rank 1 init, no adjustment to estimated cov matrices. 

```{r,fig.width=12,fig.height=4}
par(mfrow=c(1,3))

out.ed0 = simu_mash_lfsr(Ulist,pi = rep(1/length(Ulist),length(Ulist)),
                N = n*length(Ulist),adjust = NULL,
                nreps=20,mc.cores = 4,use.U.true = FALSE,seed=seed,npc=2)

plot.out(out.ed0,titles = "ED estimated cov mat, unadjusted")
```



```{r,include=FALSE,eval=FALSE}
# 2. Use udr + random unconstrained init, no adjustment to estimated cov matrices.

out.udr0 = simu_mash_lfsr_udr(Ulist,pi = rep(1/length(Ulist),length(Ulist)),
                N = n*length(Ulist),adjust = NULL,
                nreps=20,mc.cores = 4,seed=seed)

plot.out(out.udr0,titles = "TEEM estimated cov mat, unadjusted")
```

Now adjust the estimated covariance matrices from ED by adding $\sigma^2 I$, and we vary $\sigma^2$ from small to large.

```{r,fig.width=12,fig.height=4}

s_list = c(0.001,0.01,0.05,0.1,0.3,0.5,1,5,10,100)

par(mfrow=c(1,3))

for(s in s_list){
  out = simu_mash_lfsr_search(Ulist,pi = rep(1/length(Ulist),length(Ulist)),
                N = n*length(Ulist),
                s=s,
                nreps=20,mc.cores = 4,seed=seed,npc=2)
  plot.out(out,titles = paste("ED estimated cov mat, adjusted with s^2=",s))
}


```

## example 2

Try another example

```{r}
u = toeplitz(c(1,-0.7,0.2,0.1,0))
u[4:5,] = 0
u[,4:5] = 0
u
Ulist = list(U3 = u)
```


1. Using ED + rank 1 init, no adjustment to estimated cov matrices. 

```{r,fig.width=12,fig.height=4}
par(mfrow=c(1,3))

out.ed0 = simu_mash_lfsr(Ulist,pi = rep(1/length(Ulist),length(Ulist)),
                N = n*length(Ulist),adjust = NULL,
                nreps=20,mc.cores = 4,use.U.true = FALSE,seed=seed,npc=2)

plot.out(out.ed0,titles = "ED estimated cov mat, unadjusted")
```


Now adjust the estimated covariance matrices from ED by adding $\sigma^2 I$, and we vary $\sigma^2$ from small to large.

```{r,fig.width=12,fig.height=4}

s_list = c(0.001,0.01,0.05,0.1,0.3,0.5,1,5,10,100)

par(mfrow=c(1,3))

for(s in s_list){
  out = simu_mash_lfsr_search(Ulist,pi = rep(1/length(Ulist),length(Ulist)),
                N = n*length(Ulist),
                s=s,
                nreps=20,mc.cores = 4,seed=seed,npc=2)
  plot.out(out,titles = paste("ED estimated cov mat, adjusted with s^2=",s))
}

```




