---
title: "ebci discussion"
author: "Dongyue Xie"
date: "2022-02-09"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction


We perform simulation studies to compare Empirikos and ash, illustrating our discussion on choosing one global worst-case prior for selecting a subset of findings. 

We compare total five Empirikos AMARI confidence intervals in two simulation settings with ash, a total number 10 comparisons. 

The five Empirikos methods are:

1. Empirikos Gaussian scale mixture $SN(0.1,15.6,1.1)$;

2. Empirikos Gaussian location mixture $LN([-4:0.05:4],0.25)$;

3. Add a small scale mixture component with sd $10^{-5}$ to Empirikos Gaussian scale mixture $SN(0.1,15.6,1.1)$;

4. Add a point-mass $\delta_0$ to Empirikos Gaussian scale mixture $SN(0.1,15.6,1.1)$;

5. NPMLE, by using $LN([-4:0.02:4],0)$.

For the two simulation settings, we generate effects $\mu$ from 

1. $G_{spiky} = ...$;

2. $0.5\delta_0 + 0.5G_{spiky}$.

## Simulation setting 1

When running ash, we consider nullweights 1,10,100.

```{r,eval=FALSE}
library(ashr)
target = seq(-6,0,by=0.2)
nullweights = c(1,10,100)
nreps = 100
n = 5000

a = 0.1
mixsd = 0.1
while(a<15.6){
  a = a*1.1
  mixsd = c(mixsd,a)
}

g_true = normalmix(c(0.5,0.2,0.1,0.1,0.1),rep(0,5),s=c(0,0.25,0.5,1,2))
lfsr = array(dim=c(nreps,length(target),length(nullweights)))
lfsr0 = array(dim=c(nreps,length(target),length(nullweights)))
lfsr_true = ash(target,s=1,g=g_true,fixg=TRUE)$result$lfsr

set.seed(12345)
for(i in 1:nreps){
  if(i%%5==0){
    print(i)
  }
  z = read.csv(paste('output/ebci_discussion/Z0/Zs_rep',i,'.csv',sep=''))$Column1
  for(j in 1:length(nullweights)){
    ghat = ash(z,s=1,mixcompdist = 'normal',pointmass = TRUE, optmethod='mixSQP',
               prior = 'nullbiased',mixsd = mixsd,nullweight=nullweights[j])$fitted_g
    res = ash(target,s=1,g=ghat,fixg=TRUE)
    lfsr0[i,,j] = res$result$lfsr
    
    ghat = ash(z,s=1,mixcompdist = 'normal',pointmass = FALSE, optmethod='mixSQP',
               prior = 'nullbiased',mixsd = mixsd,nullweight=nullweights[j])$fitted_g
    res = ash(target,s=1,g=ghat,fixg=TRUE)
    lfsr[i,,j] = res$result$lfsr
  }
}

saveRDS(list(lfsr=lfsr,lfsr0=lfsr0,lfsr_true=lfsr_true,
             g_true=g_true,nullweights=nullweights),
        file='output/ebci_discussion/ash_Z0.rds')
```


```{r}
ash_res = readRDS('output/ebci_discussion/ash_Z0.rds')
ash_lfsr = apply(ash_res$lfsr0,3,colMeans)
colnames(ash_lfsr) = ash_res$nullweights
lfsr_true = ash_res$lfsr_true
```

```{r}
ebci = read.csv('output/ebci_discussion/spiky0.csv')
unique(ebci$method)
```

```{r}
ploter = function(method,ebci,ash_lfsr,lfsr_true,idx,target,color = rgb(0, 0, 0.8,0.4)){
  out = ebci[ebci$method == method,]
  plot(target[idx],lfsr_true[idx],type='l',xlab='z',
       ylab=expression(paste('P(',mu>=0,'|',Z==z,')')),
       ylim=c(0,0.5))
  polygon(c(rev(out$t[idx]), out$t[idx]), c(rev(out$upper_mean[idx]), out$lower_mean[idx]), col=color, border = NA)
  lines(target[idx],lfsr_true[idx])
  for(j in 1:3){
    lines(target[idx],ash_lfsr[idx,j],lty=j+1)
  }
  abline(h=0.01,lty=3,col='grey40')
  abline(h=0.05,lty=3,col='grey40')
  legend('topleft',c('Ground truth',paste('nullweight = ',ash_res$nullweights),method),
         lty=c(1,2,3,4,NA),bty="n",
         fill=c(rep(NA,4), color),
         border = rep(NA,5),
         x.intersp=c(rep(1,4),1))
}
```

```{r,eval=FALSE}
mtd_list = c("amari_scalemix","amari_scalemix_eps","amari_scalemix0", "amari_locmix",   "amari_locmix_npmle")

pdf(file = "output/ebci_discussion/spiky0_a4_23.pdf",  
    width = 12, 
    height = 8) 

par(mfrow=c(2,3))
for(method in mtd_list){
  ploter(method,ebci,ash_lfsr,lfsr_true,idx=1:16,target=seq(-6,0,by=0.2))
}

dev.off()
#knitr::include_graphics("output/ebci_discussion/spiky0_a4.pdf")
```


![alt text here]("output/ebci_discussion/spiky0_a4.pdf")




