---
title: "Empirical Bayes Confidence Interval"
author: "Dongyue Xie"
date: "2022-01-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 12,fig.height = 4)
```

## Introduction

In this simulation, we examine the EBCI method proposed by Ignatiadis and Wager(2021). And compare the results with ash.

The simulation setting is: at each of $nreps=100$ repetitions, $n=5000$ samples $z_i$ are generate from the distribution $z_i|\mu_i\sim N(\mu_i,1), \mu_i\sim g(\cdot)$, where the prior $g$ is referred to as spiky,\[g^{spiky} = 0.4N(0,0.25^2)+0.2N(0,0.5^2)+0.2N(0,1)+0.2N(0,2^2).\]

Given a class of non-parametric priors, specifically, Gaussian location mixture and Gaussian scale mixture priors, the prior $g$ is estimated based on $5000$ samples. Then we obtain the $\widehat {lfsr}(z)$(for simplicity, $p(\mu\geq 0|Z=z)$) at a grid of target $z$ values, $-5:0.2:5$. At each repetition, we get the lower and upper bound of the confidence interval, as well as the true lfsr based on true $g$. The final plot is based the average of them over all repetitions.

The EBCI package Empirikos is implemented in Julia, here I saved the output as a CSV file and imported here.

```{r}
ebci = read.csv('output/ebci/lfsr_no_point_mass/spiky_lfsr.csv')
head(ebci)
unique(ebci$method)
```

Load ash results. For code running ash experiments, see the appendix.

```{r}
ash_res = readRDS('output/ebci/ash_lfsr_no_point_mass.rds')
ash_posprob_mean = apply(ash_res$PosProb,3,colMeans)
colnames(ash_posprob_mean) = ash_res$nullweights
```

## Scale mixture


We first look at using Gaussian scale mixture as the prior. 

The prior class is $g\in SN(0.1,15.6,1.1)$.

```{r}
#t, method, ground_truth_mean,   lower_mean,   upper_mean
# "dkw_scalemix"   "kde_scalemix"   "amari_scalemix" "dkw_locmix"     "kde_locmix"     "amari_locmix"  
target = seq(-5,0,by = 0.2)
method_list = c("dkw_scalemix","kde_scalemix","amari_scalemix")


plot_res = function(method,color,idx,ylim=c(0,1)){
  out = ebci[ebci$method == method,]
  plot(out$t[idx],out$ground_truth_mean[idx],type='l',xlab='z',ylab=expression(paste('P(',mu>=0,'|',Z==z,')')),
       ylim=ylim)
  #axis(1,at=seq(-5,5,by=1))
  #axis(2,at=seq(0,1,by=0.1))
  polygon(c(rev(out$t[idx]), out$t[idx]), c(rev(out$upper_mean[idx]), out$lower_mean[idx]), col=color, border = NA)
  lines(out$t[idx],out$ground_truth_mean[idx])
  #lines(out$t,out$lower_mean,lty=2)
  #lines(out$t,out$upper_mean,lty=2)
  for(j in 1:3){
    lines(target[idx],ash_posprob_mean[idx,j],lty=j+1)
  }
  abline(h=0.01,lty=3,col='grey40')
  abline(h=0.05,lty=3,col='grey40')
  legend('topleft',c('Ground truth',paste('nullweight = ',ash_res$nullweights),method),
         lty=c(1,2,3,4,NA),bty="n",
         fill=c(rep(NA,4), color),
         border = rep(NA,5),
         x.intersp=c(rep(1,4),1))

}

#par(mfrow=c(1,3))
#plot_res(method = "dkw_scalemix",idx=1:51,color = rgb(1, 0.6, 0,0.5))
#plot_res(method = "kde_scalemix",idx=1:51,color = rgb(0.4, 0, 0,0.4))
#plot_res(method = "amari_scalemix",idx=1:51,color = rgb(0, 0, 0.8,0.4))

```



```{r,eval=FALSE,include=FALSE}
#Now let's zoom in a little bit.
#Ignore the positive z part.
par(mfrow=c(1,3))
plot_res(method = "dkw_scalemix",idx=1:26,color = rgb(1, 0.6, 0,0.5),ylim=c(0,0.5))
plot_res(method = "kde_scalemix",idx=1:26,color = rgb(0.4, 0, 0,0.4),ylim=c(0,0.5))
plot_res(method = "amari_scalemix",idx=1:26,color = rgb(0, 0, 0.8,0.4),ylim=c(0,0.5))
```

### ash and ebci

Push z to left, up to -3.

```{r}
par(mfrow=c(1,3))
plot_res(method = "dkw_scalemix",idx=1:11,color = rgb(1, 0.6, 0,0.5),ylim=c(0,0.08))
plot_res(method = "kde_scalemix",idx=1:11,color = rgb(0.4, 0, 0,0.4),ylim=c(0,0.08))
plot_res(method = "amari_scalemix",idx=1:11,color = rgb(0, 0, 0.8,0.4),ylim=c(0,0.08))
```

Further push z to left, up to -4.

```{r}
par(mfrow=c(1,3))
plot_res(method = "dkw_scalemix",idx=1:6,color = rgb(1, 0.6, 0,0.5),ylim=c(0,0.011))
plot_res(method = "kde_scalemix",idx=1:6,color = rgb(0.4, 0, 0,0.4),ylim=c(0,0.011))
plot_res(method = "amari_scalemix",idx=1:6,color = rgb(0, 0, 0.8,0.4),ylim=c(0,0.011))
```



## Location mixture

Now let's take a look at the location mixture prior class. The prior class is $LN(0.25^2,[-6,6])$. ([-6:0.05:6])
```{r,eval=FALSE,include=FALSE}
par(mfrow=c(1,3))
plot_res(method = "dkw_locmix",idx=1:51,color = rgb(1, 0.6, 0,0.5))
plot_res(method = "kde_locmix",idx=1:51,color = rgb(0.4, 0, 0,0.4))
plot_res(method = "amari_locmix",idx=1:51,color = rgb(0, 0, 0.8,0.4))

#Now let's zoom in a little bit.
#Ignore the positive z part.

par(mfrow=c(1,3))
plot_res(method = "dkw_locmix",idx=1:26,color = rgb(1, 0.6, 0,0.5),ylim=c(0,0.5))
plot_res(method = "kde_locmix",idx=1:26,color = rgb(0.4, 0, 0,0.4),ylim=c(0,0.5))
plot_res(method = "amari_locmix",idx=1:26,color = rgb(0, 0, 0.8,0.4),ylim=c(0,0.5))
```

### ash and ebci

Push z to left, up to -3.

```{r}
par(mfrow=c(1,3))
plot_res(method = "dkw_locmix",idx=1:11,color = rgb(1, 0.6, 0,0.5),ylim=c(0,0.08))
plot_res(method = "kde_locmix",idx=1:11,color = rgb(0.4, 0, 0,0.4),ylim=c(0,0.08))
plot_res(method = "amari_locmix",idx=1:11,color = rgb(0, 0, 0.8,0.4),ylim=c(0,0.08))
```

Further push z to left, up to -4.

```{r}
par(mfrow=c(1,3))
plot_res(method = "dkw_locmix",idx=1:6,color = rgb(1, 0.6, 0,0.5),ylim=c(0,0.011))
plot_res(method = "kde_locmix",idx=1:6,color = rgb(0.4, 0, 0,0.4),ylim=c(0,0.011))
plot_res(method = "amari_locmix",idx=1:6,color = rgb(0, 0, 0.8,0.4),ylim=c(0,0.011))
```


## Compare Scale and Location mixture

Overlay upper bound of scale and location mixture. And see how they diff.


```{r}
par(mfrow=c(1,3))
idx = 1:26
scalemix = ebci[ebci$method == 'dkw_scalemix',]
locmix = ebci[ebci$method == 'dkw_locmix',]
plot(locmix$t[idx],locmix$ground_truth_mean[idx],type='l',xlab='z',ylab=expression(paste('P(',mu>=0,'|',Z==z,')')),
       ylim=c(0,1))
  
lines(scalemix$t[idx],scalemix$upper_mean[idx],lty=2)
lines(locmix$t[idx],locmix$upper_mean[idx],lty=3)
legend('topleft',c('Ground truth','dkw_scalemix','dkw_locmix'),lty=c(1,2,3))


idx = 1:26
scalemix = ebci[ebci$method == 'kde_scalemix',]
locmix = ebci[ebci$method == 'kde_locmix',]
plot(locmix$t[idx],locmix$ground_truth_mean[idx],type='l',xlab='z',ylab=expression(paste('P(',mu>=0,'|',Z==z,')')),
       ylim=c(0,1))
  
lines(scalemix$t[idx],scalemix$upper_mean[idx],lty=2)
lines(locmix$t[idx],locmix$upper_mean[idx],lty=3)
legend('topleft',c('Ground truth','kde_scalemix','kde_locmix'),lty=c(1,2,3))


idx = 1:26
scalemix = ebci[ebci$method == 'amari_scalemix',]
locmix = ebci[ebci$method == 'amari_locmix',]
plot(locmix$t[idx],locmix$ground_truth_mean[idx],type='l',xlab='z',ylab=expression(paste('P(',mu>=0,'|',Z==z,')')),
       ylim=c(0,1))
  
lines(scalemix$t[idx],scalemix$upper_mean[idx],lty=2)
lines(locmix$t[idx],locmix$upper_mean[idx],lty=3)
legend('topleft',c('Ground truth','amari_scalemix','amari_locmix'),lty=c(1,2,3))
```

Again, let's zoom in the plot a little bit.

```{r}
par(mfrow=c(1,3))
idx = 1:16
scalemix = ebci[ebci$method == 'dkw_scalemix',]
locmix = ebci[ebci$method == 'dkw_locmix',]
plot(locmix$t[idx],locmix$ground_truth_mean[idx],type='l',xlab='z',ylab=expression(paste('P(',mu>=0,'|',Z==z,')')),
       ylim=c(0,1))
  
lines(scalemix$t[idx],scalemix$upper_mean[idx],lty=2)
lines(locmix$t[idx],locmix$upper_mean[idx],lty=3)
legend('topleft',c('Ground truth','dkw_scalemix','dkw_locmix'),lty=c(1,2,3))


idx = 1:16
scalemix = ebci[ebci$method == 'kde_scalemix',]
locmix = ebci[ebci$method == 'kde_locmix',]
plot(locmix$t[idx],locmix$ground_truth_mean[idx],type='l',xlab='z',ylab=expression(paste('P(',mu>=0,'|',Z==z,')')),
       ylim=c(0,1))
  
lines(scalemix$t[idx],scalemix$upper_mean[idx],lty=2)
lines(locmix$t[idx],locmix$upper_mean[idx],lty=3)
legend('topleft',c('Ground truth','kde_scalemix','kde_locmix'),lty=c(1,2,3))


idx = 1:16
scalemix = ebci[ebci$method == 'amari_scalemix',]
locmix = ebci[ebci$method == 'amari_locmix',]
plot(locmix$t[idx],locmix$ground_truth_mean[idx],type='l',xlab='z',ylab=expression(paste('P(',mu>=0,'|',Z==z,')')),
       ylim=c(0,1))
  
lines(scalemix$t[idx],scalemix$upper_mean[idx],lty=2)
lines(locmix$t[idx],locmix$upper_mean[idx],lty=3)
legend('topleft',c('Ground truth','amari_scalemix','amari_locmix'),lty=c(1,2,3))
```




