---
title: "Estimate Prior with masked z scores"
author: "Dongyue Xie"
date: "2021-01-15"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Estimate prior covariance matrices with masked z scores and run masked mash.

 - run deconvolution using masked z scores, estimate prior covairance matrices
 - run masked mash.

```{r}
library(devtools)
load_all('code/mashr/')
source('code/genData.R')
source('code/maskedmash.R')
source('code/ebupdate.R')
source('code/maskedmd.R')
source('code/utils.R')
source('code/calc_post.R')




u1 = c(1,1,0,0,0)
u2 = c(0,0,1,1,1)
cov1 = tcrossprod(u1)
cov1[1,1] = 1
cov2 = tcrossprod(u2)
cov2[5,5] = 2
cov2[4,4] = 1

U = list(U1=cov1,U2=cov2)
U
signal_sd = 2

set.seed(12345)
simdata = genData(300,U,signal_sd)
data = mash_set_data(simdata$Bhat,simdata$Shat)
```

Let's try to run mash: the ED intializaiton are some random postive definite matrices generated from Wishart distribution.

```{r}
U.c = cov_canonical(data)
# init at random
U.init = list(U1 = rWishart(1,5,diag(5))[,,1], 
              U2 = rWishart(1,5,diag(5))[,,1],
              U3 = rWishart(1,5,diag(5))[,,1])

U.init = normalize_Ulist(U.init)

U.ed = cov_ed(data,Ulist_init = U.init)
lapply(U.ed,round,digits=2)
U.ed = lapply(U.ed,function(x){x+0.01*diag(5)})
out_mash = mash(data, c(U.c,U.ed),verbose = F)
lfsr_list = seq(0.01,0.5,length.out = 100)
power_mash = c()
fdp_mash = c()
for(l in 1:length(lfsr_list)){
 temp = fdp_power(out_mash$result$lfsr,simdata$B,alpha = lfsr_list[l])
 power_mash[l] = temp$power
 fdp_mash[l] = temp$fdp
}


# init at pca
# U.pca = cov_pca(data,5,subset=NULL)
# out.ed = cov_ed(data,Ulist_init = U.pca)
# out_mash1 = mash(data, c(U.c,out.ed),verbose = F)
# fdp_power(out_mash1$result$lfsr,simdata$B)
```

Using rank one matrix needs additional attention - add the random effect. For now, we do not use rank one matrix as initialization(so the output is not rank 1).



Let's check the masked.md function by using the original z scores. The estimated ones are very similar to `cov_ed` outputs.

```{r}
out.md.check = masked.md(data,thresh=1e5,U.data = U.init,max_iter = 100,printevery = 50)
plot(out.md.check$loglik,type='l',ylab = 'loglik',xlab='iterations')
lapply(out.md.check$U.est,round,digits=2)
```

## mask all z scores

Now run "masked deconvolution".

```{r}
out.md = masked.md(data,U.data = U.init,max_iter = 100,printevery = 50)
plot(out.md$loglik,type='l',ylab = 'loglik',xlab='iterations')
lapply(out.md$U.est,round,digits=2)
round(out.md$pi,2)
out.maskedmash = masked.mash(data,U.canon = U.c,U.data = out.md$U.est,
                             U.update = 'none',fixg = F,
                             usepointmass = TRUE,verbose = TRUE,printevery = 50)

#out_mash_use_masked.md = mash(data, c(U.c,out.md$U.est),verbose = F,outputlevel = 3)
#fdp_power(out_mash_use_masked.md$result$lfsr,simdata$B)
power_maskedmash = c()
fdp_maskedmash = c()
for(l in 1:length(lfsr_list)){
 temp = fdp_power(out.maskedmash$lfsr,simdata$B,alpha = lfsr_list[l])
 power_maskedmash[l] = temp$power
 fdp_maskedmash[l] = temp$fdp
}
plot(lfsr_list,fdp_maskedmash,type='l',ylab="fdp",xlab="lfsr")
lines(lfsr_list,fdp_mash,col=3)
legend("bottomright",c("maskedmash","mash"),lty=c(1,1),col=c(1,3))
plot(lfsr_list,power_maskedmash,type='l',ylab="power",xlab="lfsr")
lines(lfsr_list,power_mash,col=3)
legend("bottomright",c("maskedmash","mash"),lty=c(1,1),col=c(1,3))
plot(fdp_maskedmash,power_maskedmash,type = 'l',ylab="power",xlab="fdp")
lines(fdp_mash,power_mash,col=3)
legend("bottomright",c("maskedmash","mash"),lty=c(1,1),col=c(1,3))
```

1. run mash using masked deconvolution output

```{r}
out_mash.md = mash(data, c(U.c,out.md$U.est),verbose = F)
power_mash.md = c()
fdp_mash.md = c()
for(l in 1:length(lfsr_list)){
 temp = fdp_power(out_mash.md$result$lfsr,simdata$B,alpha = lfsr_list[l])
 power_mash.md[l] = temp$power
 fdp_mash.md[l] = temp$fdp
}
```

2. run maskedmash using ed output

```{r}
out.maskedmash.ed = masked.mash(data,U.canon = U.c,U.data = U.ed,
                             U.update = 'none',fixg = F,
                             usepointmass = TRUE,verbose = TRUE,printevery = 50)
power_maskedmash.ed = c()
fdp_maskedmash.ed = c()
for(l in 1:length(lfsr_list)){
 temp = fdp_power(out.maskedmash.ed$lfsr,simdata$B,alpha = lfsr_list[l])
 power_maskedmash.ed[l] = temp$power
 fdp_maskedmash.ed[l] = temp$fdp
}

plot(lfsr_list,fdp_maskedmash,type='l',ylab="fdp",xlab="lfsr")
lines(lfsr_list,fdp_mash,col=3)
lines(lfsr_list,fdp_mash.md,col=2)
lines(lfsr_list,fdp_maskedmash.ed,col=4)
legend('bottomright',c('mask.mash+mask.ed',"mash+ed","mash+mask.ed","mask.mash+ed"),lty=c(1,1,1,1),col=c(1,3,2,4))
```

It seems that the masked ed needs to extra attention.

```{r}
unlist(lapply(out.md$U.est,kappa))
unlist(lapply(U.ed,kappa))
```

The condition number is huge. Let's add a diagonal matrix to stablize it.

```{r}
U.masked.re = lapply(out.md$U.est,function(x){x+0.01*diag(5)})
unlist(lapply(U.masked.re,kappa))
```

Now re-run maskedmash

```{r}
out.maskedmash.re = masked.mash(data,U.canon = U.c,U.data = U.masked.re,
                             U.update = 'none',fixg = F,
                             usepointmass = TRUE,verbose = TRUE,printevery = 50)

#out_mash_use_masked.md = mash(data, c(U.c,out.md$U.est),verbose = F,outputlevel = 3)
#fdp_power(out_mash_use_masked.md$result$lfsr,simdata$B)
power_maskedmash.re = c()
fdp_maskedmash.re = c()
for(l in 1:length(lfsr_list)){
 temp = fdp_power(out.maskedmash.re$lfsr,simdata$B,alpha = lfsr_list[l])
 power_maskedmash.re[l] = temp$power
 fdp_maskedmash.re[l] = temp$fdp
}
plot(lfsr_list,fdp_maskedmash.re,type='l',ylab="fdp",xlab="lfsr")
lines(lfsr_list,fdp_mash,col=3)
legend("bottomright",c("maskedmash","mash"),lty=c(1,1),col=c(1,3))
plot(lfsr_list,power_maskedmash.re,type='l',ylab="power",xlab="lfsr")
lines(lfsr_list,power_mash,col=3)
legend("bottomright",c("maskedmash","mash"),lty=c(1,1),col=c(1,3))
plot(fdp_maskedmash.re,power_maskedmash.re,type = 'l',ylab="power",xlab="fdp")
lines(fdp_mash,power_mash,col=3)
legend("bottomright",c("maskedmash","mash"),lty=c(1,1),col=c(1,3))
```


## masked partial z scores

```{r}
out.md2 = masked.md(data,U.data = U.init,max_iter = 100,printevery = 50,thresh = 2)
plot(out.md2$loglik,type='l',ylab = 'loglik',xlab='iterations')
lapply(out.md2$U.est,round,digits=2)
round(out.md$pi,2)
unlist(lapply(out.md2$U.est,kappa))
U.masked.re2 = lapply(out.md2$U.est,function(x){x+0.01*diag(5)})
unlist(lapply(U.masked.re2,kappa))
out.maskedmash2 = masked.mash(data,U.canon = U.c,U.data = U.masked.re2,thresh=2,
                             U.update = 'none',fixg = F,
                             usepointmass = TRUE,verbose = TRUE,printevery = 50)
out.maskedmash2$maskedProp

power_maskedmash.re2 = c()
fdp_maskedmash.re2 = c()
for(l in 1:length(lfsr_list)){
 temp = fdp_power(out.maskedmash2$lfsr,simdata$B,alpha = lfsr_list[l])
 power_maskedmash.re2[l] = temp$power
 fdp_maskedmash.re2[l] = temp$fdp
}

plot(lfsr_list,fdp_maskedmash.re,type='l',ylab="fdp",xlab="lfsr")
lines(lfsr_list,fdp_mash,col=3)
lines(lfsr_list,fdp_maskedmash.re2,col=4)
legend("bottomright",c("maskedmash","mash","maskedmash25.8%"),lty=c(1,1,1),col=c(1,3,4))


plot(lfsr_list,power_maskedmash.re,type='l',ylab="power",xlab="lfsr")
lines(lfsr_list,power_mash,col=3)
lines(lfsr_list,power_maskedmash.re2,col=4)
legend("bottomright",c("maskedmash","mash","maskedmash25.8%"),lty=c(1,1,1),col=c(1,3,4))

plot(fdp_maskedmash.re,power_maskedmash.re,type = 'l',ylab="power",xlab="fdp")
lines(fdp_mash,power_mash,col=3)
lines(fdp_maskedmash.re2,power_maskedmash.re2,col=4)
legend("bottomright",c("maskedmash","mash","maskedmash25.8%"),lty=c(1,1,1),col=c(1,3,4))
```

## Non-normal Distribution

Consider a situation when the model is mis-speicifed, for example test statistics follow a multivariate t distribution. 
