---
title: "lfsr, pointmass"
author: "Dongyue Xie"
date: "2021-04-07"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Study the effect of not using point mass in model fitting, on lfsr.

The true model is

\[z\sim \pi_0 N(0,1) + \pi_1 N(0,1+\sigma^2_1),\]

while we fit the following model

\[z\sim \pi_0N(0,\sigma^2_0+1)+\pi_1 N(0,1+\sigma^2_1),\]

where $\sigma^2_0$ is a small positive number. 

Define $\tilde{\pi}_0 = \frac{\pi_0 N(z;0,1)}{\pi_0 N(z;0,1) + \pi_1 N(z;0,1+\sigma^2_1)}$ and $\hat{\tilde{\pi}}_0 = \frac{\pi_0 N(z;0,1+\sigma^2_0)}{\pi_0N(z;0,\sigma^2_0+1)+\pi_1 N(z;0,1+\sigma^2_1)}$.

```{r}
library(ashr)
# z ~ pi0*N(0,1+s0) + pi1*N(0,1+s1) 
lfsr1 = function(z,pi0,pi1,s0,s1){

  pm1 = s1/(s1+1)*z
  pv1 = s1/(s1+1)
  np1 = pnorm(0,pm1,sqrt(pv1))
  
  pi0_tilde = pi0*dnorm(z,0,sd = sqrt(1+s0))/(pi0*dnorm(z,0,sd = sqrt(1+s0)) + pi1*dnorm(z,0,sd = sqrt(1+s1)))
  pi1_tilde = 1-pi0_tilde
  if(s0==0){
    
    np = pi0_tilde + pi1_tilde*np1
    pp = pi0_tilde + pi1_tilde*(1-np1)
  }else{
    pm0 = s0/(s0+1)*z
    pv0 = s0/(s0+1)
    np0 = pnorm(0,pm0,sqrt(pv0))
    np = pi0_tilde*np0 + pi1_tilde*np1
    pp = pi0_tilde*(1-np0) + pi1_tilde*(1-np1)
  }
  lfsr = pmin(np,pp)
  
  return(list(lfsr=lfsr,np=np,pi0_tilde = pi0_tilde))
}
```

We vary $z$ from 0 to 5, and plot $\hat{\tilde{\pi}}_0$ vs $\tilde{\pi}_0$, and $\widehat{lfsr}$ vs $lfsr$.

In this example, we set $\sigma^2_1 = 4$ and $\sigma^2_0 = 0.01$, $\pi_0=0.8$.


```{r}
n = 1000
z = seq(0,5,length.out = n)
pi0 = 0.8
pi1 = 1-pi0
s0 = 0.01
s1 = 4
fit.true = lfsr1(z,pi0,pi1,0,s1)
fit.est = lfsr1(z,pi0,pi1,s0,s1)

plot(z,fit.true$pi0_tilde,type = 'l',ylab = 'pi0_tilde')
lines(z,fit.est$pi0_tilde,type='l',col=4)
legend("topright",c('true','estimated'),col=c(1,4),lty=c(1,1))

plot(z,log(fit.true$pi0_tilde)-log(fit.est$pi0_tilde),type='l',ylab = 'log(pi0_tilde/pi0_tilde_hat)')
abline(h=(3/4*pi1/pi0*s0/sqrt(1+s1)),col = 4,lty=2)
#lines(z[-(1:(n/2))],s0/2*(1-z[-(1:(n/2))]^2)*(1-pi0/pi1*sqrt(1+s1)*exp(-0.5*s1/(1+s1)*z[-(1:(n/2))]^2)),col=4)
lines(z,s0/2*(1-z^2)*(1-pi0/pi1*sqrt(1+s1)*exp(-0.5*s1/(1+s1)*z^2)),col=4)
legend('bottomleft',c('true','approx when z small', "appox when z large"),col=c(1,4,4),lty=c(1,2,1))
```

When $z$ is small, $\hat{\tilde{\pi}}_0$ is very close to the true one, while as $z$ gets larger,it always over-estimate the true one.

```{r}
plot(z,fit.true$lfsr,type = 'l',ylab = 'lfsr')
lines(z,fit.est$lfsr,type='l',col=4)
legend("topright",c('true','estimated'),col=c(1,4),lty=c(1,1))

plot(z[900:1000],fit.true$lfsr[900:1000],type = 'l',ylab = 'lfsr',xlab='z',ylim = range(c(fit.true$lfsr[900:1000],fit.est$lfsr[900:1000])))
lines(z[900:1000],fit.est$lfsr[900:1000],type='l',col=4)
legend("topright",c('true','estimated'),col=c(1,4),lty=c(1,1))

plot(fit.true$lfsr,fit.est$lfsr,type='l',xlab='True lfsr', ylab="Estimated lfsr")
abline(a=0,b=1,lty=2)

plot(z,fit.true$lfsr/fit.est$lfsr,type='l',ylab = "lfsr / lfsr_hat")

plot(z,fit.true$lfsr-fit.est$lfsr,type='l',ylab = "lfsr-lfsr_hat")

plot(fit.true$lfsr-fit.est$lfsr, fit.true$pi0_tilde*(1-pnorm(-sqrt(s0/(1+s0))*z)),type='l',xlab = 'lfsr - lfsr_hat', ylab = 'estimate of (lfsr - lfsr_hat)')
abline(a=0,b=1,lty=2)

plot(z,log(fit.true$lfsr-fit.est$lfsr),type='l',ylab = "log(lfsr-lfsr_hat)")
```

$\widehat{lfsr}$ always under-estimates $lfsr$.

We now set $\sigma^2_0 = 0.1$ and re-run the analysis

```{r}
n = 1000
z = seq(0,5,length.out = n)
pi0 = 0.8
pi1 = 1-pi0
s0 = 0.1
s1 = 4
fit.true = lfsr1(z,pi0,pi1,0,s1)
fit.est = lfsr1(z,pi0,pi1,s0,s1)

plot(z,fit.true$pi0_tilde,type = 'l',ylab = 'pi0_tilde')
lines(z,fit.est$pi0_tilde,type='l',col=4)
legend("topright",c('true','estimated'),col=c(1,4),lty=c(1,1))

plot(z,log(fit.true$pi0_tilde)-log(fit.est$pi0_tilde),type='l',ylab = 'log(pi0_tilde/pi0_tilde_hat)')
abline(h=(3/4*pi1/pi0*s0/sqrt(1+s1)),col = 4,lty=2)
#lines(z[-(1:(n/2))],s0/2*(1-z[-(1:(n/2))]^2)*(1-pi0/pi1*sqrt(1+s1)*exp(-0.5*s1/(1+s1)*z[-(1:(n/2))]^2)),col=4)
lines(z,s0/2*(1-z^2)*(1-pi0/pi1*sqrt(1+s1)*exp(-0.5*s1/(1+s1)*z^2)),col=4)
legend('bottomleft',c('true','approx when z small', "appox when z large"),col=c(1,4,4),lty=c(1,2,1))
```


```{r}
plot(z,fit.true$lfsr,type = 'l',ylab = 'lfsr')
lines(z,fit.est$lfsr,type='l',col=4)
legend("topright",c('true','estimated'),col=c(1,4),lty=c(1,1))

plot(z[900:1000],fit.true$lfsr[900:1000],type = 'l',ylab = 'lfsr',xlab='z',ylim = range(c(fit.true$lfsr[900:1000],fit.est$lfsr[900:1000])))
lines(z[900:1000],fit.est$lfsr[900:1000],type='l',col=4)
legend("topright",c('true','estimated'),col=c(1,4),lty=c(1,1))

plot(fit.true$lfsr,fit.est$lfsr,type='l',xlab='True lfsr', ylab="Estimated lfsr")
abline(a=0,b=1,lty=2)

plot(z,fit.true$lfsr/fit.est$lfsr,type='l',ylab = "lfsr / lfsr_hat")

plot(z,fit.true$lfsr-fit.est$lfsr,type='l',ylab = "lfsr-lfsr_hat")

plot(fit.true$lfsr-fit.est$lfsr, fit.true$pi0_tilde*(1-pnorm(-sqrt(s0/(1+s0))*z)),type='l',xlab = 'lfsr - lfsr_hat', ylab = 'estimate of (lfsr - lfsr_hat)')
abline(a=0,b=1,lty=2)

plot(z,log(fit.true$lfsr-fit.est$lfsr),type='l',ylab = "log(lfsr-lfsr_hat)")
```

Let's try an example using ash.

```{r}
n = 10000
pi0 = 0.8
pi1 = 1-pi0
s0 = 0
s1 = 4
set.seed(12345)
z = c(rnorm(n*pi0),rnorm(n*(1-pi0),0,sqrt(s1+1)))
#hist(z,breaks = 100)

fit.true = lfsr1(z,pi0,pi1,0,s1)

out1 = ash(z,1,"normal",pointmass=TRUE,gridmult = 1.3)
out1$fitted_g

plot(fit.true$lfsr,out1$result$lfsr,pch='.',ylab='lfsr ash pointmass',xlab='lfsr')
abline(a=0,b=1,lty=2,col='grey50')

out2 = ash(z,1,'normal',pointmass=FALSE,gridmult = 1.3)
out2$fitted_g
plot(fit.true$lfsr,out2$result$lfsr,pch='.',ylab='lfsr ash no pointmass',xlab='lfsr')
abline(a=0,b=1,lty=2,col='grey50')

plot(z,fit.true$lfsr - out2$result$lfsr,pch='.',ylab = "lfsr - lfsr ash no pointmass")
```

How about we swap things - generate z from a model without pointmass but we fit a model with point mass.

```{r}
set.seed(12345)
s0 = 0.1
z = c(rnorm(n*pi0,0,sqrt(1+s0)),rnorm(n*(1-pi0),0,sqrt(s1+1)))
#hist(z,breaks = 100)

fit.true = lfsr1(z,pi0,pi1,s0,s1)

out1 = ash(z,1,"normal",pointmass=TRUE,gridmult = 1.2,nullweight=10)
out1$fitted_g

plot(fit.true$lfsr,out1$result$lfsr,pch='.',ylab='lfsr ash pointmass',xlab='lfsr')
abline(a=0,b=1,lty=2,col='grey50')

out2 = ash(z,1,'normal',pointmass=FALSE,gridmult = 1.2,nullweight=10)
out2$fitted_g
plot(fit.true$lfsr,out2$result$lfsr,pch='.',ylab='lfsr ash no pointmass',xlab='lfsr')
abline(a=0,b=1,lty=2,col='grey50')

out3 = ash(z,1,'normal',pointmass=FALSE,gridmult = 1.2,nullweight=1)
out3$fitted_g
plot(fit.true$lfsr,out3$result$lfsr,pch='.',ylab='lfsr ash no pointmass',xlab='lfsr')
abline(a=0,b=1,lty=2,col='grey50')


```



## Conclusion

If the true model has a point mass while we fit a model without it, then the estimated lfsr under-estimates the true lfsr. 







