---
title: "On lfsr monotonicity"
author: "Dongyue Xie"
date: "2022-02-09"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Overview

We study if lfsr is monotonic in $z$ and/or $|z|$.

## NPMLE

We focus on likelihood $z|\mu\sim N(0,1)$.

Let's start with the prior class NPMLE, which in practice is a mixture of pointmass across a dense grid,

$$g^{NPMLE} = \sum_k\pi_k\delta(m_k).$$
The dense grid is usually chosen to span the range of $z$, from example $-5:0.02:5$. The default number of grid is 300 in `GLmix()` from `REBayes` package while in `ebnm`, it suggests 100 is enough.

Given $\boldsymbol\pi$ and $\boldsymbol m$, the posterior of $\mu$ is 

$$\mu|z\sim \sum_k\tilde\pi_k\delta(m_k), \text{ where } \tilde\pi_k = \frac{\pi_kN(z;m_k,1)}{\sum_l\pi_lN(z;m_l,1)}.$$

The lfsr can be obtained by calculating $$p(\mu\leq 0|z) = \sum_{k:m_k\leq 0}\tilde\pi_k, \text{ and }p(\mu\geq 0|z) = \sum_{k:m_k\geq 0}\tilde\pi_k.$$


```{r}
lfsr_npmle = function(z,w,m){
  n = length(z)
  lfsr = c()
  for(i in 1:n){
    zi = z[i]
    w_tilde_i = w*dnorm(zi,mean = m,sd = 1)
    w_tilde_i = w_tilde_i/sum(w_tilde_i)
    lfsr[i] = min(sum(w_tilde_i[m<=0]),sum(w_tilde_i[m>=0]))
  }
  lfsr
}
```

Symmetric means Symmetric about 0.

### Symmetric mean

Weights uniform

```{r}
K = 101
z = seq(-5,5,by = 0.2)
w = rep(1/K,K)
m = seq(-5,5,length.out=K)
plot(m,w,type='l',xlab='mean',ylab='weight')
abline(v = 0,lty=2,col='grey80')
plot(z,lfsr_npmle(z,w,m),type='l',xlab='z',ylab='lfsr')
abline(v = 0,lty=2,col='grey80')
```

Weights not uniform, but symmetric about 0.

```{r}
K = 101
m = seq(-5,5,length.out=K)
z = seq(-5,5,by = 0.2)
w = c(1:floor(K/2),2*K,floor(K/2):1)
w = w/sum(w)
plot(m,w,type='l',xlab='mean',ylab='weight')
abline(v = 0,lty=2,col='grey80')
plot(z,lfsr_npmle(z,w,m),type='l',xlab='z',ylab='lfsr')
abline(v = 0,lty=2,col='grey80')
```

Weights not symmetric about 0.

```{r}
K = 101
m = seq(-5,5,length.out=K)
z = seq(-5,5,by = 0.2)
w = c(1:floor(K/2+1),2*K,floor(K/2-1):1)
w = w/sum(w)
plot(m,w,type='l',xlab='mean',ylab='weight')
abline(v = 0,lty=2,col='grey80')
plot(z,lfsr_npmle(z,w,m),type='l',xlab='z',ylab='lfsr')
abline(v = 0,lty=2,col='grey80')
```


Weights not symmetric about 0.

```{r}
K = 101
m = seq(-5,5,length.out=K)
z = seq(-5,5,by = 0.2)
w = c(1:floor(K/2+10),2*K,floor(K/2-10):1)
w = w/sum(w)
plot(m,w,type='l',xlab='mean',ylab='weight')
abline(v = 0,lty=2,col='grey80')
plot(z,lfsr_npmle(z,w,m),type='l',xlab='z',ylab='lfsr')
abline(v = 0,lty=2,col='grey80')
```


Weights symmetric about 0 but not uni-modal.

```{r}
K = 101
m = seq(-5,5,length.out=K)
z = seq(-5,5,by = 0.1)
w = c(1:(floor((K-1)/4)-1),5*K,floor((K-1)/4):1,10*K,1:(floor((K-1)/4)),5*K,(floor((K-1)/4)-1):1)
w = w/sum(w)
plot(m,w,type='l',xlab='mean',ylab='weight')
abline(v = 0,lty=2,col='grey80')
plot(z,lfsr_npmle(z,w,m),type='l',xlab='z',ylab='lfsr')
abline(v = 0,lty=2,col='grey80')
```

```{r}
K = 101
m = seq(-5,5,length.out=K)
z = seq(-5,5,by = 0.1)
w = c(1:(floor((K-1)/4)-1),5*K,floor((K-1)/4):1,1,1:(floor((K-1)/4)),5*K,(floor((K-1)/4)-1):1)
w = w/sum(w)
plot(m,w,type='l',xlab='mean',ylab='weight')
abline(v = 0,lty=2,col='grey80')
plot(z,lfsr_npmle(z,w,m),type='l',xlab='z',ylab='lfsr')
abline(v = 0,lty=2,col='grey80')
```


### Not symmetric mean

mean skew to the right

```{r}
K = 101
z = seq(-5,5,by = 0.1)
w = rep(1/K,K)
m = seq(-5,6,length.out=K)
plot(m,w,type='l',xlab='mean',ylab='weight')
abline(v = 0,lty=2,col='grey80')
plot(z,lfsr_npmle(z,w,m),type='l',xlab='z',ylab='lfsr')
abline(v = 0,lty=2,col='grey80')
```



```{r}
K = 101
z = seq(-5,5,by = 0.1)
w = c(1:floor(K/2),2*K,floor(K/2):1)
w = w/sum(w)
m = seq(-1,4,length.out=K)
plot(m,w,type='l',xlab='mean',ylab='weight')
abline(v = 0,lty=2,col='grey80')
plot(z,lfsr_npmle(z,w,m),type='l',xlab='z',ylab='lfsr')
abline(v = 0,lty=2,col='grey80')
```

```{r}
K = 105
z = seq(-5,5,by = 0.1)
w = c(1:40,20*K,64:1)
w = w/sum(w)
m = seq(-5,8,length.out=K)
plot(m,w,type='l',xlab='mean',ylab='weight')
abline(v = 0,lty=2,col='grey80')
plot(z,lfsr_npmle(z,w,m),type='l',xlab='z',ylab='lfsr')
abline(v = 0,lty=2,col='grey80')
```


```{r}
K = 105
z = seq(-5,5,by = 0.1)
w = c(1:40,20*K,64:30,20*K,28:1)
w = w/sum(w)
m = seq(-5,8,length.out=K)
plot(m,w,type='l',xlab='mean',ylab='weight')
abline(v = 0,lty=2,col='grey80')
plot(z,lfsr_npmle(z,w,m),type='l',xlab='z',ylab='lfsr')
abline(v = 0,lty=2,col='grey80')
```





