---
title: "masked mash on FDR control, t and uniform prior"
author: "Dongyue Xie"
date: "2021-03-10"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

```{r}
library(devtools)
library(parallel)
load_all('code/mashr/')
files.sources = list.files('code/maskedmashr/')
sapply(files.sources, function(x){source(paste('code/maskedmashr/',x,sep=''))})
#'@title get summary from simu study
#'@description given a sequence of
get.simu.result = function(out,alpha){

  fdr_result = mclapply(out,function(x){
    
    non_null_idx = which(x$data$B!=0)
    rej.mash = mashFDR(x$mash.out,alpha)
    #fp.mash = c(fsp(x$data$Bhat[rej.mash],x$data$B[rej.mash]),
    #            powr(rej.mash,non_null_idx))
    fp.mash = c(fdp(rej.mash,non_null_idx),powr(rej.mash,non_null_idx))
    fp = lapply(x[-c(1:2)],function(z){
      rej = maskedmashFDR(x$data$Bhat,z,alpha)$rej.set
      fdps = fdp(rej,non_null_idx)
      powers = powr(rej,non_null_idx)
      c(fdps,powers)
    })

    cbind(fp.mash,do.call(cbind,fp))
  },mc.cores = 4)

  temp = do.call(rbind,fdr_result)
  fdr.idx = seq(1,2*length(out),2)
  list(FDP = temp[fdr.idx,],POWER = temp[-fdr.idx,])
}

plot.fdr = function(out,
                    alpha_list = seq(0,0.3,length.out = 30),
                    method_list = c("mash","mash.on.maskedZ.all",
                                    "masked.mash.all","masked.mash.70%",
                                    "masked.mash.40%","masked.mash.20%")){
  n_method = length(method_list)
  fdr = c()
  powers = c()
  for(i in 1:length(alpha_list)){
    temp = get.simu.result(out,alpha_list[i])
    fdr = rbind(fdr,colMeans(temp$FDP))
    powers = rbind(powers,colMeans(temp$POWER))
  }
  
  plot(alpha_list,fdr[,1],type='l',lwd=2,ylim = range(fdr),xlab = 'Target FDR Level', ylab = 'FDR')
  for(i in 2:ncol(fdr)){
    lines(alpha_list,fdr[,i],lwd=2,col=i,lty=i)
  }
  abline(a=0,b=1,lty=1,col='grey80')
  legend('bottomright',method_list,
         lty = 1:n_method,col=1:n_method,lwd=rep(2,n_method))
  
  plot(alpha_list,powers[,1],type='l',lwd=2,ylim = range(powers),xlab = 'Target FDR Level', ylab = 'Power')
  for(i in 2:ncol(powers)){
    lines(alpha_list,powers[,i],lwd=2,col=i,lty=i)
  }
  legend('bottomright',method_list,
         lty = 1:n_method,col=1:n_method,lwd=rep(2,n_method))
  
  
  plot(fdr[,1],powers[,1],type='l',lwd=2,xlim = range(fdr),ylim = range(powers),xlab = 'FDR', ylab = 'Power')
  for(i in 2:ncol(powers)){
    lines(fdr[,i],powers[,i],lwd=2,col=i,lty=i)
  }
  legend('bottomright',method_list,
         lty = 1:n_method,col=1:n_method,lwd=rep(2,n_method))
  
  
}


```

## t-prior

Generate 2000 samples from 4 multivariate t mixtures with df = 5, and covariance matrix: null, identity,
```{r}
u = toeplitz(c(1,0.4,0.3,0.2,0.1))
u[3:5,] = 0
u[,3:5] = 0
u

u2 = toeplitz(c(1,-0.4,0.3,0.2,0.1))
u2[1:3,] = 0
u2[,1:3] = 0
u2
```


Repeat the process 20 times.  


```{r}
out = readRDS('output/maskedmashFDR/t_prior_SNR2K4N2000df5.rds')
plot.fdr(out)
```

## uniform distribution

Generate 2000 samples from 4 multivariate uniform distribution(marginally [-a,a], obtained from transforming multivariate normals)

```{r}
out = readRDS('output/maskedmashFDR/uniform_prior_SNR2K4N2000.rds')
plot.fdr(out)
```


## half uniform distribution

Generate 2000 samples from 4 multivariate uniform distribution(marginally [0,a]).

```{r}
out = readRDS('output/maskedmashFDR/halfuniform_prior_SNR2K4N2000.rds')
plot.fdr(out)
```
