---
title: "masked mash on FDR control, t distribution"
author: "Dongyue Xie"
date: "2021-02-24"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Consider a model-misspecification case: $z_{ir}|\mu_{ir}\sim t_\nu(\mu_{ir})$, where $t_\nu(\mu_{ir})$ is a non-central t-distribution with df $\nu$ and non-central parameter $\mu_{ir}$. If $\mu_{ir} = 0$, then it reduces to a t distribution with df $\nu$. Parameters $\mu$ are generated from mixture of normals, 

$$\mu_i\sim 0.25N(0,0)+0.25N(0,3\times I) + 0.25N(0,3\times U_1) + 0.25N(0,3\times U_2).$$

There are two ways to fit a masked mash model on the dataset. One is to directly fit the model on the given dataset; another one is to transform the p values(from t distribution) to "z scores" by $sign(x)*\Phi^{-1}(1-p/2)$, then fit a model on the "z scores". The latter way is more similar to AdaPT, that starts with p values and choose a working model. And the working model is masked mash in this context.

We generate 2000 samples from 4 Normal mixtures with covariance matrix: null, identity, equal effects in first 2 conditions, equal effects in the last 3 conditions. Repeat the process 30 times.  

```{r,message=FALSE,results=FALSE}
library(parallel)
library(devtools)
load_all('code/mashr/')
files.sources = list.files('code/maskedmashr/')
sapply(files.sources, function(x){source(paste('code/maskedmashr/',x,sep=''))})

#'@title get summary from simu study
#'@description given a sequence of
get.simu.result = function(out,alpha){

  fdr_result = mclapply(out,function(x){
    #print(length(x))
    non_null_idx = which(x$data$B!=0)
    #print(non_null_idx)
    fp = lapply(x[-1],function(z){
      rej = mashFDR(x$data$Bhat,z,alpha)$rej.set
      fdps = fdp(rej,non_null_idx)
      powers = powr(rej,non_null_idx)
      c(fdps,powers)
    })

    do.call(cbind,fp)
  },mc.cores = 4)

  temp = do.call(rbind,fdr_result)
  fdr.idx = seq(1,2*length(out),2)
  list(FDP = temp[fdr.idx,],POWER = temp[-fdr.idx,])
}
```

## Direct fit on test statisitcs

```{r}
out = readRDS('output/maskedmashFDR/t_SNR3K4N2000.rds')

alpha_list = seq(0,0.5,length.out = 50)
fdr = c()
powers = c()
for(i in 1:length(alpha_list)){
  temp = get.simu.result(out,alpha_list[i])
  fdr = rbind(fdr,colMeans(temp$FDP))
  powers = rbind(powers,colMeans(temp$POWER))
}

#########

plot(alpha_list,fdr[,1],type='l',lwd=2,ylim = range(fdr),xlab = 'Target FDR Level', ylab = 'FDR')
for(i in 2:ncol(fdr)){
  lines(alpha_list,fdr[,i],lwd=2,col=i,lty=i)
}
abline(a=0,b=1,lty=1,col='grey80')
legend('bottomright',c("mash","mash.on.maskedZ.all","mash.on.maskedZ.40%","masked.mash.all","masked.mash.40%"),
       lty = 1:5,col=1:5,lwd=rep(2,5))

#########

plot(alpha_list,powers[,1],type='l',lwd=2,ylim = range(powers),xlab = 'Target FDR Level', ylab = 'Power')
for(i in 2:ncol(powers)){
  lines(alpha_list,powers[,i],lwd=2,col=i,lty=i)
}
legend('bottomright',c("mash","mash.on.maskedZ.all","mash.on.maskedZ.40%","masked.mash.all","masked.mash.40%"),
       lty = 1:5,col=1:5,lwd=rep(2,5))

#########

plot(fdr[,1],powers[,1],type='l',lwd=2,xlim = range(fdr),ylim = range(powers),xlab = 'FDR', ylab = 'Power')
for(i in 2:ncol(powers)){
  lines(fdr[,i],powers[,i],lwd=2,col=i,lty=i)
}
legend('bottomright',c("mash","mash.on.maskedZ.all","mash.on.maskedZ.40%","masked.mash.all","masked.mash.40%"),
       lty = 1:5,col=1:5,lwd=rep(2,5))
```

## Fit model on transformed p values

```{r}
#'@title get summary from simu study
#'@description given a sequence of
get.simu.resultP = function(out,alpha){

  fdr_result = mclapply(out,function(x){
    #print(length(x))
    non_null_idx = which(x$data$B!=0)
    #print(non_null_idx)
    fp = lapply(x[-1],function(z){
      rej = mashFDR(x$data$Bhat,z,alpha,x$data$P)$rej.set
      fdps = fdp(rej,non_null_idx)
      powers = powr(rej,non_null_idx)
      c(fdps,powers)
    })

    do.call(cbind,fp)
  },mc.cores = 4)

  temp = do.call(rbind,fdr_result)
  fdr.idx = seq(1,2*length(out),2)
  list(FDP = temp[fdr.idx,],POWER = temp[-fdr.idx,])
}
```

```{r}
out = readRDS('output/maskedmashFDR/t_P_SNR3K4N2000.rds')

alpha_list = seq(0,0.5,length.out = 50)
fdr = c()
powers = c()
for(i in 1:length(alpha_list)){
  temp = get.simu.resultP(out,alpha_list[i])
  fdr = rbind(fdr,colMeans(temp$FDP))
  powers = rbind(powers,colMeans(temp$POWER))
}

#########

plot(alpha_list,fdr[,1],type='l',lwd=2,ylim = range(fdr),xlab = 'Target FDR Level', ylab = 'FDR')
for(i in 2:ncol(fdr)){
  lines(alpha_list,fdr[,i],lwd=2,col=i,lty=i)
}
abline(a=0,b=1,lty=1,col='grey80')
legend('bottomright',c("mash","mash.on.maskedZ.all","mash.on.maskedZ.40%","masked.mash.all","masked.mash.40%"),
       lty = 1:5,col=1:5,lwd=rep(2,5))

#########

plot(alpha_list,powers[,1],type='l',lwd=2,ylim = range(powers),xlab = 'Target FDR Level', ylab = 'Power')
for(i in 2:ncol(powers)){
  lines(alpha_list,powers[,i],lwd=2,col=i,lty=i)
}
legend('bottomright',c("mash","mash.on.maskedZ.all","mash.on.maskedZ.40%","masked.mash.all","masked.mash.40%"),
       lty = 1:5,col=1:5,lwd=rep(2,5))

#########

plot(fdr[,1],powers[,1],type='l',lwd=2,xlim = range(fdr),ylim = range(powers),xlab = 'FDR', ylab = 'Power')
for(i in 2:ncol(powers)){
  lines(fdr[,i],powers[,i],lwd=2,col=i,lty=i)
}
legend('bottomright',c("mash","mash.on.maskedZ.all","mash.on.maskedZ.40%","masked.mash.all","masked.mash.40%"),
       lty = 1:5,col=1:5,lwd=rep(2,5))
```

